### 2024-05-25 paper/figs/definitions.tex
some work on model diagram for paper

\usepackage{xcolor}

% Color definitions
\definecolor{customorange}{rgb}{0.94, 0.55, 0}
\definecolor{customteal}{rgb}{0.47, 0.52, 0.6}
\definecolor{customgreen}{rgb}{0.18, 0.62, 0.27}
\definecolor{fontcolor}{rgb}{0.204, 0.227, 0.251}

% Tikz shape definitions
\tikzstyle{nn-box} = [rectangle, rounded corners=0.5mm, minimum width=35mm, minimum height=10mm, text centered, draw=customteal, fill=customteal!30, text=fontcolor]
\tikzstyle{transformer-box} = [rectangle, rounded corners=0.5mm, minimum width=35mm, minimum height=10mm, text centered, draw=customgreen, fill=customgreen!30, text=fontcolor]
\tikzstyle{xgb-box} = [rectangle, rounded corners=0.5mm, minimum width=35mm, minimum height=10mm, text centered, draw=customorange, fill=customorange!30, text=fontcolor]
\tikzstyle{pca-box} = [rectangle, rounded corners=0.5mm, minimum width=20mm, minimum height=7mm, text centered, draw=customorange!70, fill=customorange!10, text=fontcolor]

% \tikzstyle{startstop} = [rectangle, rounded corners, minimum width=3cm, minimum height=1cm,text centered, draw=black, fill=red!30]
% \tikzstyle{process} = [rectangle, minimum width=3cm, minimum height=1cm, text centered, draw=black, fill=orange!30]
% \tikzstyle{arrow} = [thick,->,>=stealth]