### 2024-05-25 paper/main.tex

\documentclass{article}
% \usepackage[utf8]{inputenc}
\usepackage{xcolor}
\usepackage{tikz}
\usetikzlibrary{shapes.geometric, arrows}

% Color definitions
\definecolor{customorange}{rgb}{0.94, 0.55, 0}
\definecolor{customteal}{rgb}{0.47, 0.52, 0.6}
\definecolor{customgreen}{rgb}{0.18, 0.62, 0.27}
\definecolor{fontcolor}{rgb}{0.204, 0.227, 0.251}

\author{
    Rawlings, Darren\\
    \texttt{email@domain.loc}      % If needed
    \and 
    Chopard, Timothy\\
    \texttt{email@domain.loc}
}

\title{An Inspiring and Groundbreaking Look at Plants}

\begin{document}
    \maketitle
    \include{sections/01-intro}
    \include{figs/model-diagram}
\end{document}