### 2024-05-25 paper/main.tex
refactored the latex document into sections in subdirectories, added to diagram, updated some details

%% The first command in your LaTeX source must be the \documentclass command.
%%
%% Options:
%% twocolumn : Two column layout.
%% hf: enable header and footer.
\documentclass[
% twocolumn,
% hf,
]{ceurart}

\usepackage{tikz}

%%
%% One can fix some overfulls
\sloppy

%%
%% Minted listings support 
%% Need pygment <http://pygments.org/> <http://pypi.python.org/pypi/Pygments>
\usepackage{minted}
%% auto break lines
\setminted{breaklines=true}

% Color definitions
\definecolor{customorange}{rgb}{0.94, 0.55, 0}
\definecolor{customteal}{rgb}{0.47, 0.52, 0.6}
\definecolor{customgreen}{rgb}{0.18, 0.62, 0.27}
\definecolor{fontcolor}{rgb}{0.204, 0.227, 0.251}
%%
%% end of the preamble, start of the body of the document source.
\begin{document}

\input{sections/00-title-abstract}
\input{sections/01-intro}
%%%%%

\input{sections/99-template.tex}
\end{document}