### 2024-05-25 paper/sections/01-intro.tex

\section{Introduction}

\begin{figure}
    \begin{center}
        \input{figs/model-diagram.tex}
        \caption{This diagram is just here temporarily and will be in a more relevant section when complete}
    \end{center}
\end{figure}