### 2024-05-28 paper/sections/04-method.tex
initial methods structure written

\clearpage  % TODO comment this out when finished

\section{Method}

In order to manage the multilabel classification required for this research we implemented an ensemble approach. 
This section details the individual architectures and the methods used to combine them.

\subsection{Architectures}

The core architectures used in this project were 18 layer ResNet, XGBoost and [TRANSFORMER]. The outputs of these were 
then weighted and combined before the maximum arguments were selected.

\subsubsection{XGBoost}
XGBoost is an open source gradient tree boosting package \cite{xgboost}. For this research we used the xgbregression model It has shown broad success on a number of 
\subsubsection{[Resnet]}
\subsubsection{[Transformer]}

\subsection{Process}
\subsubsection{Species scoring}
\subsubsection{Species counts}
\subsection{Metrics}